Reliable error estimation for Sobol' indices\thanks{This work is supported by the CITiES project funded by the Agence Nationale de la Recherche (grant ANR-12-MONU-0020) and by the United States National Science Foundation (grant DMS-1522687).}

### 2016-10-02 article/article_for_submission.tex
changes L, PCA + comm fig1

Then, as an alternative to the common Sobol' indices estimators (\ref{common.sobol.est}), we can define the following two estimators with their respective error bounds:
\begin{gather}
\begin{aligned}
\widehat{\underline{S}}_u & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \underline{S}_u(\vI),1\Big{)} + \max\Big{(}\min_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \underline{S}_u(\vI),0\Big{)} \right) \\
\varepsilon_{\widehat{\underline{S}}_u} & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \underline{S}_u(\vI),1\Big{)} - \max\Big{(}\min_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \underline{S}_u(\vI),0\Big{)} \right)
\end{aligned}
\label{formula_undersu}
\end{gather}
and,
\begin{gather}
\begin{aligned}
\widehat{\overline{S}}_u & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \overline{S}_u(\vI),1\Big{)} + \max\Big{(}\min_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \overline{S}_u(\vI),0\Big{)} \right) \\
\varepsilon_{\widehat{\overline{S}}_u} & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \overline{S}_u(\vI),1\Big{)} - \max\Big{(}\min_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \overline{S}_u(\vI),0\Big{)} \right)
\end{aligned}
\label{formula_oversu}
\end{gather}
Because numerator and denominator are both known to be positive, maximizing $\underline{S}_u(\vI)$ (resp. $\overline{S}_u(\vI)$) is done through maximizing the numerator $I_1$ (resp. $I_2$) and minimizing the denominator $I_3-I_4^2$. Analogously, to minimize $\underline{S}_u(\vI)$ (resp. $\overline{S}_u(\vI)$) one minimizes the numerator $I_1$ (resp. $I_2$) and maximizes the denominator $I_3-I_4^2$. As an example, Figure \ref{fig:1} illustrates the region of possible values of $\underline{S}_u(\vI)$ given the true value of $I_1$.
\begin{figure}[!ht]
\caption{The delimited region on the figure represents the values $\underline{S}_2$ in $\vI\in B_{(0,0,0.05,0.1)}(\widehat{\vI})$ for the Bratley \textit{et al.} function. There, $\underline{S}_2^{\min}=\max\Big{(}\min_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \underline{S}_2(\vI),0\Big{)}$ and $\underline{S}_2^{\max}=\min\Big{(}\max_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \underline{S}_2(\vI),1\Big{)}$, and $\widehat{\underline{S}}_2=(\underline{S}_2^{\min}+\underline{S}_2^{\max})/2$.}
\centering
\includegraphics[width=.5\textwidth]{Images/estimator_3d_small.eps}
\label{fig:1}       % Give a unique label
\end{figure}

@@

The discretized Brownian motion $\vect{B}$ follows a multivariate normal distribution with mean $\vzero$ and covariance matrix $\mathsf{\Sigma}$, with $\mathsf{\Sigma}_{ij}=\min(t_i,t_j)$. If $\vect{Y}\sim N(\vzero,I_d)$, given a matrix $M$ such that $\mathsf{\Sigma}=MM^t$, then $M\vect{Y}\sim N(\vzero,\mathsf{\Sigma})$. Hence, we can generate the discretized Brownian motion using the inverse normal cumulative distribution function $\phi^{-1}$: $\vect{B}\sim M\phi^{-1}(\vect{X})$ with $\phi^{-1}(\vect{X})=(\phi^{-1}(X_1),\dots,\phi^{-1}(X_d))^t$ and $\vect{X}\sim \mathcal{U}([0,1]^d)$. Thus, we can estimate the Sobol' indices over $f_{\rm payoff}(M\phi^{-1}(\vect{X}))$. In this particular application, the choice of $M$ is pivotal to reduce the effective dimensionality of the model. In this regard, Sobol' indices can help us assess whether we reduce the effective dimensionality. We will compare the case with either the Cholesky decomposition and the PCA construction.

@@

Table \ref{res.AsianPCA.Ab}, shows the results for Variant $A.b$ when $M$ is constructed with the PCA decomposition.

\begin{table}[!ht]
\caption{Averaged values of $\widehat{\underline{S}}_u$, ${\widehat{\overline{S}}_u}$ and total number of evaluations of Variant $A.a$ with PCA decomposition.}
\centering
\begin{tabular}{ccccc}
\hline
 input & $\widehat{\underline{S}}_u$ & $n$ & $\widehat{\overline{S}}_u$ & $n$ \\ \hline
 $x_1$ & $0.9800$ & $131072$ & $0.9960$ & $117309$ \\ \hline
 $x_2$ & $0.0037$ & $952$ & $0.0172$ & $824$ \\ \hline
 $x_3$ & $0.0016$ & $517$ & $0.0023$ & $512$ \\ \hline
 $x_4$ & $0.0014$ & $512$ & $0.0005$ & $512$ \\ \hline
 $x_5$ & $0.0005$ & $512$ & $0.0001$ & $512$ \\ \hline
 $x_6$ & $0.0002$ & $512$ & $0.0000$ & $512$ \\ \hline
 \multicolumn{3}{l}{Total number of evaluations: }
 \\ \hline
\end{tabular}
\label{res.AsianPCA.Aa}
\end{table}

@@

\begin{table}[!ht]
\caption{Averaged values of $\widehat{\underline{S}}_u$ and total number of evaluations of Variant $B$ with PCA decomposition.}
\centering
\begin{tabular}{ccc}
\hline
 input & $\widehat{\underline{S}}_u$ & $n$ \\ \hline
 $x_1$ & $0.9800$ & $131072$ \\ \hline
 $x_2$ & $0.0037$ & $18422$ \\ \hline
 $x_3$ & $0.0028$ & $7834$ \\ \hline
 $x_4$ & $0.0029$ & $10486$ \\ \hline
 $x_5$ & $0.0017$ & $4613$ \\ \hline
 $x_6$ & $0.0029$ & $9467$ \\ \hline \hline
\multicolumn{2}{l}{Total number of evaluations: } \\ \hline 
\end{tabular}
\label{res.AsianChol.B}
\end{table}

One may conclude that these estimates explain how the PCA Brownian motion construction reduces the high effective dimensionality of the payoff function, in contrast to the initial Cholesky decomposition.

>>>>

Then, as an alternative to the common Sobol' indices estimators (\ref{common.sobol.est}), we can define the following two estimators with their respective error bounds:
\begin{gather}
\begin{aligned}
\widehat{\underline{S}}_u & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \underline{S}_u(\vI),1\Big{)} + \max\Big{(}\min_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \underline{S}_u(\vI),0\Big{)} \right) \\
\varepsilon_{\widehat{\underline{S}}_u} & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \underline{S}_u(\vI),1\Big{)} - \max\Big{(}\min_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \underline{S}_u(\vI),0\Big{)} \right)
\end{aligned}
\label{formula_undersu}
\end{gather}
and,
\begin{gather}
\begin{aligned}
\widehat{\overline{S}}_u & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \overline{S}_u(\vI),1\Big{)} + \max\Big{(}\min_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \overline{S}_u(\vI),0\Big{)} \right) \\
\varepsilon_{\widehat{\overline{S}}_u} & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \overline{S}_u(\vI),1\Big{)} - \max\Big{(}\min_{\vI\in B_{\varepsilon_{\widehat{\vI}}}(\widehat{\vI})} \overline{S}_u(\vI),0\Big{)} \right)
\end{aligned}
\label{formula_oversu}
\end{gather}
Because numerator and denominator are both known to be positive, maximizing $\underline{S}_u(\vI)$ (resp. $\overline{S}_u(\vI)$) is done through maximizing the numerator $I_1$ (resp. $I_2$) and minimizing the denominator $I_3-I_4^2$. Analogously, to minimize $\underline{S}_u(\vI)$ (resp. $\overline{S}_u(\vI)$) one minimizes the numerator $I_1$ (resp. $I_2$) and maximizes the denominator $I_3-I_4^2$. As an example, Figure \ref{fig:1} illustrates the region of possible values of $\underline{S}_2(\vI)$ given the true value of $I_1$ for the test function of Bratley \textit{et al.} \cite{Bratley}. This function is further described in section \ref{sec:5}. 
\begin{figure}[!ht]
\caption{The delimited region on the figure represents the values $\underline{S}_2$ in $\vI\in B_{(0,0,0.05,0.1)}(\widehat{\vI})$ for the Bratley \textit{et al.} function. There, $\underline{S}_2^{\min}=\max\Big{(}\min_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \underline{S}_2(\vI),0\Big{)}$ and $\underline{S}_2^{\max}=\min\Big{(}\max_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \underline{S}_2(\vI),1\Big{)}$, and $\widehat{\underline{S}}_2=(\underline{S}_2^{\min}+\underline{S}_2^{\max})/2$.}
\centering
\includegraphics[width=.5\textwidth]{Images/estimator_3d_small.eps}
\label{fig:1}       % Give a unique label
\end{figure}

@@

The discretized Brownian motion $\vect{B}$ follows a multivariate normal distribution with mean $\vzero$ and covariance matrix $\mathsf{\Sigma}$, with $\mathsf{\Sigma}_{ij}=\min(t_i,t_j)$. If $\vect{Y}\sim N(\vzero,I_d)$, given a matrix $M$ such that $\mathsf{\Sigma}=MM^t$, then $M\vect{Y}\sim N(\vzero,\mathsf{\Sigma})$. Hence, we can generate the discretized Brownian motion using the inverse normal cumulative distribution function $\phi^{-1}$: $\vect{B}\sim M\phi^{-1}(\vect{X})$ with $\phi^{-1}(\vect{X})=(\phi^{-1}(X_1),\dots,\phi^{-1}(X_d))^t$ and $\vect{X}\sim \mathcal{U}([0,1]^d)$. Thus, we can estimate the Sobol' indices over $f_{\rm payoff}(M\phi^{-1}(\vect{X}))$. In this particular application, the choice of $M$ is pivotal to reduce the effective dimensionality of the model. In this regard, Sobol' indices can help us assess whether we reduce the effective dimensionality. We will compare the case with either the Cholesky decomposition or the PCA construction.

@@

Tables \ref{res.AsianPCA.Aa}, \ref{res.AsianPCA.Ab}, and \ref{res.AsianPCA.B} show the results when $M$ corresponds to the PCA decomposition of $\mathsf{\Sigma}$.

\begin{table}[!ht]
\caption{Averaged values of $\widehat{\underline{S}}_u$, ${\widehat{\overline{S}}_u}$ and total number of evaluations of Variant $A.a$ with PCA decomposition.}
\centering
\begin{tabular}{ccccc}
\hline
 input & $\widehat{\underline{S}}_u$ & $\widehat{\overline{S}}_u$ \\ \hline
 $x_1$ & $0.9800$ & $0.9960$  \\ \hline
 $x_2$ & $0.0037$ & $0.0172$  \\ \hline
 $x_3$ & $0.0016$ & $0.0023$  \\ \hline
 $x_4$ & $0.0014$ & $0.0005$  \\ \hline
 $x_5$ & $0.0005$ & $0.0001$  \\ \hline
 $x_6$ & $0.0002$ & $0.0000$ \\ \hline
 \multicolumn{3}{l}{Total number of evaluations: $396 \ 221$}
 \\ \hline
\end{tabular}
\label{res.AsianPCA.Aa}
\end{table}

@@

\begin{table}[!ht]
\caption{Averaged values of $\widehat{\underline{S}}_u$ and total number of evaluations of Variant $B$ with PCA decomposition.}
\centering
\begin{tabular}{ccc}
\hline
 input & $\widehat{\underline{S}}_u$  \\ \hline
 $x_1$ & $0.9800$ \\ \hline
 $x_2$ & $0.0037$ \\ \hline
 $x_3$ & $0.0028$ \\ \hline
 $x_4$ & $0.0029$ \\ \hline
 $x_5$ & $0.0017$ \\ \hline
 $x_6$ & $0.0029$ \\ \hline \hline
\multicolumn{2}{l}{Total number of evaluations: $262 \ 144$} \\ \hline 
\end{tabular}
\label{res.AsianPCA.B}
\end{table}
Variant $A.b$ performs equally well as Variant $A.a$ in terms of average total number of evaluations while most probably giving better estimates for the small first-order indices. Once 	again, variant $B$ shows an improvement on the total number of evaluations with the drawback of slightly overestimating the small first-order indices.

Overall, each variant captures well how the PCA Brownian motion construction reduces the high effective dimensionality of the payoff function, in contrast to the initial Cholesky decomposition.
